---
title: "Reproducible_Research Project 1"
author: "YL Chang"
date: "Monday, May 11, 2015"
output: html_document
---

##Loading and preprocessing the data

The following code chunk reads the data and store them in 3 arrays.

```{r, echo=FALSE}
setwd("c:/users/13360/desktop/coursera/Reproducible_Research/Week_2/Project")
```
```{r}
data <- read.csv("activity.csv")
steps <-data$steps
date <- data$date
interval <-data$interval
```

##What is mean total number of steps taken per day?

Calculate the total number of steps taken per day. This is done by using the "tapply" function.

```{r}
stepdate<-tapply(steps, date, sum, na.rm = TRUE)
```

Make a histogram of the total number of steps taken each day

```{r}
hist(stepdate, main = "Total Step per Day",xlab = "Steps", col = "red")
```

This distribution is skewed to the left. This is because that there are a lot of 
NA in the original data set.

Calculate and report the mean and median of the total number of steps taken per day

```{r}
mean(stepdate)
median(stepdate)
```

##What is the average daily activity pattern?

Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) 
and the average number of steps taken, averaged across all days (y-axis)

```{r}
stepinterval<-tapply(steps, interval, mean, na.rm = TRUE)
intervallevels <- levels(factor(interval))
plot(stepinterval~intervallevels,type = "l", ylab = "number of steps", col = "red")
```

Which 5-minute interval, on average across all the days in the dataset, 
contains the maximum number of steps?

```{r}
intervallevels[which.max(stepinterval)]
```

##Imputing missing values

Calculate and report the total number of missing values in the dataset

```{r}
sum(is.na(steps))
```

Devise a strategy for filling in all of the missing values in the dataset. 
Using mean of the 5-minute intervals.

```{r}
newsteps <- steps
meanstepinterval<-tapply(newsteps, interval, mean, na.rm = TRUE)
```

Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
ind <- which( is.na( newsteps ) )
for (i in 1:length(ind)){
    k <- ind[i]%%dim(meanstepinterval)
    if(k == 0) {
      newsteps[ind[i]] <- meanstepinterval[dim(meanstepinterval)]
    } else {
      newsteps[ind[i]] <- meanstepinterval[k]
    }
}
```

Make a histogram of the total number of steps taken each day and 

```{r}
newstepdate<-tapply(newsteps, date, sum, na.rm = TRUE)
hist(newstepdate, main = "Total Step per Day",xlab = "Steps", col = "red")
```

After imputing the missing values by using the mean of the 5-minute interval accross all dates, the skewed distribution become more normal.

Calculate and report the mean and median total number of steps taken per day. 

```{r}
mean(newstepdate)
median(newstepdate)
```

##Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable in the dataset with two levels ??? ???weekday??? and ???weekend???

```{r}
weekdaydata <- weekdays(as.Date(date,"%m/%d/%Y"))
WDWE<-rep("weekday", each = length(weekdaydata))
WDWE[which(weekdaydata == "Saturday" | weekdaydata == "Sunday")] <- "weekend"
```

Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis)
and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r}
splitnewsteps<-split(newsteps,WDWE)
splitinterval<-split(interval,WDWE)

splitnewstepintervalWD<-tapply(splitnewsteps$weekday, splitinterval$weekday, mean, na.rm = TRUE)
splitnewstepintervalWE<-tapply(splitnewsteps$weekend, splitinterval$weekend, mean, na.rm = TRUE)

par(mfrow = c(1, 2))
    plot(splitnewstepintervalWD~intervallevels,type = "l",col = "red", ylab = "number of steps", main = "Weekday steps per Interval")
    plot(splitnewstepintervalWE~intervallevels,type = "l",col = "red", ylab = "number of steps", main = "Weekend steps per Interval")
```

It can be seen from the plots that this test object is more active in the weekend than in the weekday.